---
title: "Winter 2019 Math Bridge Week 1 Assignment"
author: "Henry Otuadinma"
date: "25 December 2018"
output: html_document
link:  http://rpubs.com/jackv13/454523
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mosaicCalc)
```

Find the derivatives with the respect to x of the following

1) $$F(x\mid x \ge 0)=1-e^{-\lambda x}$$


   $$\implies \frac{d}{dx}*1 - \frac {d}{dx}(-e^{-\lambda x})$$
   $$\implies \frac{d}{dx}*1 - \frac {d}{dx}(-e^{-\lambda x}) \implies 0 - \frac {d}{dx}(-e^{-\lambda x})$$
   $$\implies - (-\lambda\frac {d}{dx}x)*e^{-\lambda x} = \lambda e^{-\lambda x}$$ 
   $$\implies F(x\mid x \ge 0)=1-e^{-\lambda x} = \lambda e^{-\lambda x}$$
   
```{r}
f1 <- D(1-e^{-(lambda * x)}~x, x=10)
print(f1)
```
   

2) $$F(x\mid b>a)= \frac {x-a}{b-a}$$
   $$= \frac{d}{dx}(\frac {x-a}{b-a}) = \frac{1}{b-a}\frac{d}{dx}{x-a} = \frac{1}{b-a}(\frac{d}{dx}x - \frac{d}{dx}a)$$
   
   $$\implies F(x\mid b>a)= \frac {x-a}{b-a} = \frac{1}{b-a}$$
```{r}
f2 <- D((x-a)/(b-a)~x, x=0)
print(f2)
```
   

3) $$F(x|a<x \le c \le b)=\frac {(x-a)^2}{(b-a)(c-a)}$$

$$ =\frac {1}{(b-a)(c-a)} * \frac {d}{dx}((x-a)^2)$$

  $$u = (x-a), \implies \frac {1}{(b-a)(c-a)} * \frac {d}{du}u^2 *  \frac {d}{dx}(x-a)$$
  
  $$ = \frac {1}{(b-a)(c-a)} * 2u *  1$$
  
  $$ = \frac {2(x-a)}{(b-a)(c-a)}$$

```{r}
f3 <- D((x-a)^2/(b-a)*(c-a)~x, x=0)
print(f3)
```




4) $$ F(x|a \le c<x<b)=1-\frac {(b-x)^2}{(b-a)(c-a)}$$

$$u = (b-x), \implies \frac {d}{dx} * 1- \frac {1}{(b-a)(c-a)} * \frac {d}{du}u^2 - \frac {d}{dx}(b-x)$$
$$\implies 0 - \frac {1}{(b-a)(c-a)} * 2*(b-x) - (-1)$$

$$ = \frac {2(b-x)}{(b-a)(c-a)}$$
```{r}
f4 <- D((b-x)^2/(b-a)*(c-a)~x, x=0)
print(f4)
```


Solve the following definite and indefinite integrals

5) $$\int_0^{10}3x^3dx$$

  
   $$ = 3\int_0^{10}x^3dx = 3\frac{x^{3+1}}{3+1}$$
  
   $$ = 3\frac{x^{4}}{4} + C$$
   $$ \lim_{x \to 0} 3\frac{0^{4}}{4} = 0$$
   $$ \lim_{x \to 10} 3\frac{10^{4}}{4} = 7500$$
   
```{r}
f5 <- antiD(3*x^3~x, x=10)
print(f5)
print(f5())
```


6) $$\int \lambda xe^{-\lambda x}dx$$

   $$\implies u = -\lambda x; \space \space \lambda\int \frac {xe^u}{\lambda^2}udu = \frac {\lambda}{\lambda ^2} \int xe^{u}udu; \space \space let \int e^udu = e^u \implies$$
   $$ \frac {\lambda}{\lambda ^2} (e^uu - e^u) = \frac {\lambda}{\lambda ^2} (e^{-\lambda x}(-\lambda x) - e^{-\lambda x})$$
   $$ = \frac {\lambda}{\lambda ^2} (-\lambda e^{-\lambda x}x - e^{-\lambda x}) + C$$

```{r}
f6 <- antiD(lambda * x * e^{- (lambda * x)}~x)
print(f6)
```


7) $$\int_0^{.5}\frac{1}{b-a}dx$$

   $$\implies\int_0^{.5}\frac{1}{b-a}dx =\frac{x}{b-a} + C \implies$$
   $$\lim_{x \to 0}\frac{0}{b-a} = 0,$$
   $$lim_{x \to 0.5}\frac{0.5}{b-a} = \frac{0.5}{b-a}$$

```{r}
f7 <- antiD((1/(b-a))~x, x=.5)
print(f7)
```


8) $$ \int_0^{x}x \frac {1}{\Gamma(\alpha)\beta^\alpha}x^{\alpha - 1}e^{- \beta x}$$

With the following matrix,
```{r}
X <- matrix(c(1,3,4,2,3,6,3,1,8), 3, 3)
print(X)
```

9) Invert it using Gaussian row reduction.
```{r}
XI <- cbind(X, diag(3))
print(XI)
```
```{r}

(XI[2,] <- XI[2,] - 3*XI[1,])
print("R2 = R2 - 3R1")
print(XI)

(XI[3,] <- XI[3,] - 4*XI[1,])
print("R3 = R3 - 4R1")
print(XI)

(XI[2,] <- XI[2,] - XI[3,])
print("R2 = R2 - R3")
print(XI)

(XI[3,] <- XI[3,] - 2*XI[2,])
print("R3 = R3 - 2R2")
print(XI)

(XI[2,] <- XI[2,] + XI[3,])
print("R2 = R2 + R3")
print(XI)

(XI[3,] <- XI[3,] / 4)
print("R3 = R3 / 4")
print(XI)

(XI[1,] <- XI[1,] - 3*XI[3,])
print("R1 = R1 - 3R3")
print(XI)

(XI[2,] <- XI[2,] * (-1))
print("R2 = R2 * -1")
print(XI)

(XI[1,] <- XI[1,] - 2*XI[2,])
print("R1 = R1 - 2R2")
print(XI)

```


10) Find the determinant.
```{r}
D <- X
(D[3,] <- D[3,]/2)
print("row3 = row3/2")
print(XI)

(D[3,] <- D[3,] - D[1,])
print("row3 = row3 - row1")
print(D)

(D[2,] <- D[2,] - D[1,])
print("row2 = row2 - row1")
print(D)

(D[2,] <- D[2,] - 2*D[3,])
print("row2 = row2 - 2 * row3")
print(D)

(D[3,] <- D[1,] - D[3,])
print("row3 = row1 - row3")
print(D)

(D[3,] <- D[3,] + D[2,])
print("row3 = row3 + row2")
print(D)

detD <- D[1,1] * D[2,2] * D[3, 3]
paste("The determinant is: ", detD)
```


11) Conduct LU decomposition

Let's find the Upper triangular matrix first. Then the Lower triangular matrix will be made up of
1's at the diagonal, the multiplication coefficients used to find the upper triangular matrix, and 
0's every other places

```{r}
 XU <- X

 (XU[2,] <- XU[2,] - 3*XU[1,])
 print("row2 = row2 - 3 * row1")
 print(XU)
 
 (XU[3,] <- XU[3,] - 4*XU[1,])
 print("row3 = row3 - 4 * row1")
 print(XU)
 
 (XU[3,] <- XU[3,] - 2/3*XU[2,])
 print("row3 = row3 - 2/3 * row2")
 print(XU)
```
Therefore, the upper triangular matrix is U = 

```{r}
print(XU)
```

While L =
```{r}
L <- matrix(c(1,3,4,0,1,2/3,0,0,1), 3,3)
print(L)
```


12) Multiply the matrix by it's inverse

The inverse of the matrix is the set of numbers at the right side of the result gotten after inverting matrix X 
in question 9 as shown below

     [,1] [,2] [,3] [,4] [,5]  [,6]
[1,]    1    0    0 -4.5 -0.5  1.75
[2,]    0    1    0  5.0  1.0 -2.00
[3,]    0    0    1 -1.5 -0.5  0.75

The inverse is (-4.5, 5.0, -1.5, -0.5, 1.0, -0.5, 1.75, -2.00, 0.75)
```{r}
Inv <- matrix(c(-4.5, 5.0, -1.5, -0.5, 1.0, -0.5, 1.75, -2.00, 0.75), 3,3)
print(Inv)
```
Matrix multiplied by its inverse will therefore be X*Inv

```{r}
XInv <- X * Inv
print ("X multiplied by its inverse: ")
print(XInv)
```

