---
title: "DATA606 Homework 5"
author: "Henry Otuadinma"
date: "24 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 5.6 Working backwards, Part II. 
A 90% confidence interval for a population mean is (65, 77). The population distribution is approximately normal and the population standard deviation is unknown. This confidence interval is based on a simple random sample of 25 observations. Calculate the sample mean, the margin of error, and the sample standard deviation.

```{r}
smean= (77+65)/2

paste('The Sample mean is:', round(smean, 4))
```

We know that the margin of error ME is $\frac{x2???x1}{2}$, and we have that confidence interval CI is (x1,x2)

```{r}
n <- 25
x1 <- 65
x2 <- 77

ME <- (x2 - x1) / 2

paste('The margin of error (ME) is:', round(ME, 4))
```

For sample standard deviation, $ME = t??????SE$, We will usthe qt() function, df = 25 - 1.

```{r}
df <- 25 - 1
p <- 0.9
p2tails <- p + (1 - p)/2

tval <- qt(p2tails, df)

# Since ME = t * SE
SE <- ME / tval

# Since SE = sd/sqrt(n)
sd <- SE * sqrt(n)

paste('The standard deviation is:', round(sd, 4))
```

#### 5.14 SAT scores. 
SAT scores of students at an Ivy League college are distributed with a
standard deviation of 250 points. Two statistics students, Raina and Luke, want to estimate the
average SAT score of students at this college as part of a class project. They want their margin of
error to be no more than 25 points.
(a) Raina wants to use a 90% confidence interval. How large a sample should she collect?

Answer:

we have that $ME = z???SE$ and $SE = \frac{sd}{\sqrt(n)} => ME = Z.\frac{sd}{\sqrt(n)} => n = (\frac{z.sd}{ME})^2$

```{r}
z <- 1.65 # 90% CI
ME <- 25
sd <- 250

n <- ((z * sd) / ME ) ^ 2
paste('Raina should use a sample of size ', n)
```
From the above, the original calculation yielded 272.25 so it ia approximated to 273 students


(b) Luke wants to use a 99% confidence interval. Without calculating the actual sample size,
determine whether his sample should be larger or smaller than Raina's, and explain your
reasoning.
(c) Calculate the minimum required sample size for Luke.


Luke's sample should be larger since it will require a higher z number multiplied by the standard deviation and then squared.

(c) Calculate the minimum required sample size for Luke.

```{r}

z <- 2.575 # due to 99% Confidence interval
ME <- 25
sd <- 250

n <- ((z * sd) / ME ) ^ 2

```









