---
title: "Winter 2019 R Bridge Week 2 Assignment"
author: "Henry Otuadinma"
date: "30 December 2018"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the summary function to gain an overview of the data set. Then display the mean and
median for at least two attributes

```{r}
## the dataset used in this work is taken from the 1997-2001 British Election Panel Study (BEPS).
## A data frame with 1525 observations on the following 10 variables:
## vote, age, economic.cond.national, economic.cond.household, Blair, Hague, Kennedy, Europe, political.knowledge, gender

beps <- read.csv(file="beps.csv", header=TRUE, sep=",")
summary(beps) #Summary of the csv data
```
```{r}
## Mean, Median of age and economic.cond.national attributes

mean_age <- mean(beps$age)
paste0("Mean age: ", mean_age)

median_age <- median(beps$age)
paste0("Median age: ", median_age)

mean_national_condition <- mean(beps$economic.cond.national)
paste0("Mean Nationa Economic condition: ", mean_national_condition)

median_national_condition <- median(beps$economic.cond.national)
paste0("Median Nationa Economic condition: ", median_national_condition)
```

2. Create a new data frame with a subset of the columns and rows. Make sure to rename it.
Q3

```{r}
##creating a new data.frame with subset of the rows (selecting 20 rows) and columns of the BEPS Data
df <- subset(beps, X < 21, select = c("vote","age","economic.cond.national", "economic.cond.household", "Europe"))
df
```

3. Create new column names for the new data frame.

```{r}
names(df)<- c("voter", "age group", "national worth", "individual worth", "Euro") ## Renaming the columns

df
```

4. Use the summary function to create an overview of your new data frame. Then print the mean
and median for the same two attributes. Please compare.
```{r}
summary(df)
```

```{r}

## Mean and Median of age and economic.cond.national attributes in df (New dataframe)

mean_age <- mean(df$`age group`)
paste0("New Mean age: ",mean_age)

median_age <- median(df$`age group`)
paste0("New Median age: ", median_age)

mean_national_worth <- mean(df$`national worth`)
paste0("New Mean Nationa Economic condition: ",mean_national_worth)

median_national_condition <- median(df$`individual worth`)
paste0("New Median Nationa Economic condition: ", median_national_condition)

```

The Mean and median ages of the original data (at 54 and 53 respectively) are higher than that of 
the new dataframe (at 51 and 45 respectively). The mean ages were almost unchanged while the median ages
remained the same.


5. For at least 3 values in a column please rename so that every value in that column is renamed.
For example, suppose I have 20 values of the letter "e" in one column. Rename those values so
that all 20 would show as "excellent".

```{r}

df$voter <- gsub("^Labour$", "excellent", df$voter) ## Renaming values in the voter column

df

```

6. Display enough rows to see examples of all of steps 1-5 above.
```{r}
head(df,10) ## showing only 10 rows out of the original subsetted 20 rows
```

7. BONUS - place the original .csv in a github file and have R read from the link. This will be a very
useful skill as you progress in your data science education and career.

```{r}
beps_git <- read.csv2("https://raw.githubusercontent.com/henryvalentine/MSDS2019/master/beps.csv", header = TRUE, sep=",")
summary(beps_git) #Summary of the csv data from github.com
```

