---
title: "DATA607 Week 12 Homework: Graph Databases"
author: "Henry Otuadinma"
date: "April 22, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(DBI)
library(DT)
library(RMySQL)
library(knitr)
library(mongolite)
```

#### Problem: I am required to migrate data from a relational database like mySql to a document/graph oriented database such as Neo4J or MongoDB


##### I created a free account on <a href="https://remotemysql.com">remotemysql.com</a>, created a mySql database, created a `movies` and `ratings` tables, and then populated them with the movies and ratings data from Week 2 homework. I then created an account on <a href="https://cloud.mongodb.com">mongodb cloud</a>. I then created a collection `cunygraph` and a database `movies` and then populated it with the data from the MySQL database.


```{r echo=FALSE}
mongoCredentials <- 'jackv13:N9ufx12345'
mongoConn <- "@cunymsds2019-fm8cu.mongodb.net/cunygraph?retryWrites=true"
mongoCollection = "cunygraph"
mongoDbName = "movies"
mySqlUser <- "aey4NigbWj"
mySqlPassword <- "2A3iBxOJza"
mySqlDbName <- "aey4NigbWj"
```


##### Create the connections

```{r}

#mongodb connection string
mongoConnStr <- paste("mongodb+srv://", mongoCredentials, mongoConn, sep = "")

#mySql connection
mySqlConnStr <- dbConnect(RMySQL::MySQL(), dbname = mySqlDbName, host="remotemysql.com", port=3306, user = mySqlUser, password = mySqlPassword)

#Mongocollection
mongoCollection <- mongo(collection = mongoCollection, db = mongoDbName, url = mongoConnStr, verbose = FALSE, options = ssl_options())

```


##### Enumerate the tables in the mySql database
```{r}
dbListTables(mySqlConnStr)
```

##### pull the data in mySql database into a dataframe and create a subset
```{r}

moviesQuery <-  fetch(dbSendQuery(mySqlConnStr, "SELECT * FROM movie m
LEFT JOIN (SELECT ratedby, movieid, date, rating, `desc` FROM ratings r) r
ON m.id = r.movieid
order by m.id
  "))
# necessary because the LEFT JOIN pulls the parents and children and make the parent appear duplicated based on the number of children
movies <-distinct(moviesQuery, id, poster, title, releasedate, genre, details, summary)

# for index table
mv <- data.frame(c(movies$title), c(movies$releasedate), c(movies$genre), c(movies$details))

```

#### Print the data from mySql db with Datatable

```{r results='asis'}

datatable(mv, colnames= c("Movie", "Release Date", "Genre", "Details"), class = 'cell-border stripe', options = list(
  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#26868d', 'color': '#fff', 'text-align': 'center !important'});",
    "$(this.api().table().body()).css({'color': '#000', 'text-align': 'center !important'});",
    "}")
))

```


##### Use the `mongoCollection` created earlier to insert the data into the mongodb

```{r results='asis'}

mongoCollection$insert(mv)

```

```{r}
mongoCollection$count()

```

##### Pull all the data from the mongodb (I pulled all the data at once because they are not many: 6 in numbers)

```{r}

mongoMovies = mongoCollection$find('{}')

```


#### Print the data from the `mongoCollection` with Datatable

```{r}

datatable(mongoMovies, colnames= c("Movie", "Release Date", "Genre", "Details"), class = 'cell-border stripe', options = list(
  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#26868d', 'color': '#fff', 'text-align': 'center !important'});",
    "$(this.api().table().body()).css({'color': '#000', 'text-align': 'center !important'});",
    "}")
))

```











