---
title: "DATA607 Week7 Homework"
author: "Henry Otuadinma"
date: "16 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(rvest)
library(stringr)
library(jsonlite)
library(DT)
```

#### Working with XML and JSON in R

Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more
than one author. For each book, include the title, authors, and two or three other attributes that you find
interesting.
Take the information that you've selected about these three books, and separately create three files which
store the book's information in HTML (using an html table), XML, and JSON formats (e.g. "books.html",
"books.xml", and "books.json"). To help you better understand the different file structures, I'd prefer that you
create each of these files "by hand" unless you're already very comfortable with the file formats.
Write R code, using your packages of choice, to load the information from each of the three sources into
separate R data frames. Are the three data frames identical?

##### Solution
##### The files were created in Notepad++ first starting with json array and then the others.

```{r}
XmlB <- xmlParse(read_xml('https://raw.githubusercontent.com/henryvalentine/MSDS2019/master/Classes/DATA%20607/Home%20works/Week7/books.xml'))
bXml <- xmlRoot(XmlB)
booksData <- xmlSApply(bXml,function(x) xmlSApply(x, xmlValue))
booksXml <- data.frame(t(booksData),row.names=NULL)
booksXml
```

#### books.json

```{r}
booksJson <- read_json("https://raw.githubusercontent.com/henryvalentine/MSDS2019/master/Classes/DATA%20607/Home%20works/Week7/books.json", simplifyVector = TRUE)
datatable(mixedBooks, colnames= c("Title", "Author", "Genre", "Rating", "Number of reviewers", "Type", "Price($)"), class = 'cell-border stripe', options = list(
  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#1f77b4', 'color': '#fff', 'text-align': 'center !important'});",
    "$(this.api().table().body()).css({'color': '#000', 'text-align': 'center !important'});",
    "}")
))
```

```{r}
booksHtml<- as.data.frame(html_nodes(read_html('https://raw.githubusercontent.com/henryvalentine/MSDS2019/master/Classes/DATA%20607/Home%20works/Week7/books.html'), "table")%>% html_table(fill=TRUE))

datatable(booksHtml, colnames= c("Title", "Type", "Author(s)", "Genre", "Date Publish", "Type", "Rating", "Number of Reviewers"), class = 'cell-border stripe', options = list(
  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#26868d', 'color': '#fff', 'text-align': 'center !important'});",
    "$(this.api().table().body()).css({'color': '#000', 'text-align': 'center !important'});",
    "}")
))
```


#### books.html

```{r}
booksHtml<- as.data.frame(html_nodes(read_html('https://raw.githubusercontent.com/henryvalentine/MSDS2019/master/Classes/DATA%20607/Home%20works/Week7/books.html'), "table")%>% html_table(fill=TRUE))

datatable(booksHtml, colnames= c("Title", "Type", "Author(s)", "Genre", "Date Publish", "Type", "Rating", "Number of Reviewers"), class = 'cell-border stripe', options = list(
  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#26868d', 'color': '#fff', 'text-align': 'center !important'});",
    "$(this.api().table().body()).css({'color': '#000', 'text-align': 'center !important'});",
    "}")
))
```

##### This needs to be cleaned up

```{r}

htmlBooks <- booksHtml%>% dplyr::filter(!(title==""))

datatable(htmlBooks, colnames= c("Title", "Type", "Author(s)", "Genre", "Date Publish", "Type", "Rating", "Number of Reviewers"), class = 'cell-border stripe', options = list(
  initComplete = JS(
    "function(settings, json) {",
    "$(this.api().table().header()).css({'background-color': '#26868d', 'color': '#fff', 'text-align': 'center !important'});",
    "$(this.api().table().body()).css({'color': '#000', 'text-align': 'center !important'});",
    "}")
))
```

The dataframes are similar except that reading from the xml file reordered the column headers alphabetically. The Html file has some trailing spaces which had to be cleaned up.