---
title: "DATA605 Homework 2"
author: "Henry Otuadinma"
date: "08/09/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Problem set 1
(1) Show that AT*A != A*AT in general. (Proof and demonstration.)

```{r}

(A <- matrix(c(3,2,1,4,3,1,-4,2,-5), nrow=3, byrow=TRUE))

```

```{r}
#A Transpose
(At <- t(A))

```


```{r}

A%*%At

```


```{r}

At%*%A

```


### To check for the given condition using logical evaluation
```{r}

A %*% At == At %*% A

```
#### This shows that A*At != At*A


### (2) For a special type of square matrix A, we get AT*A = A*AT . Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).

### Answer
#### We get AT*A = A*AT for a square matrix A if A is symmetrical along the diagonal. This implies that multiplying a matrix A by its transpose AT is like multiplying the matrix by itself.

for instance:

```{r}

(A <- matrix(c(1,2,0,2,1,0,0,0,1), nrow=3, byrow=TRUE))

```

AT will then be

```{r}

(At <- t(A))

```

Checking the statement above:

```{r}

A == At

```


```{r}

A %*% At == At %*% A

```


### 2. Problem set 2
#Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track fights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your fight using radars. Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer

#### Solution:
Choosing the LU version => 

```{r}
luDecomposer <- function(A) 
{
  # if A is not a square matrix, return null
  if (dim(A)[1]!=dim(A)[2]) 
  {
    return(NULL)
  }
  
  M <- A
  dimOfA <- dim(A)[1]
  diagA <- diag(dimOfA)
  
  if (dimOfA == 1) 
  {
    return(list(diagA, M))
  }
  
  #Using a nested for-loop
  for(a in 2:dimOfA)
  {
    for(b in 1:(a-1)) 
    {
      mx <- (-M[a,b] / M[b,b]) #devicing a multiplier
      M[a, ] <- mx * M[b, ] + M[a, ]
      diagA[a,b] <- -mx
    }
  }
  return(list(diagA,M))
}

```


#### Verifications

```{r}

(A <- matrix(c(4,1,1,1,1,2,5,4,3), nrow=3, byrow=TRUE))

```


```{r}

luD <- luDecomposer(A)

```


```{r}
(l<-luD[[1]])

```

```{r}

(u<-luD[[2]])

```

How did it fare?

```{r}
A == l %*% u

```



