---
title: "DATA605 Week3 Homework"
author: "Henry Otuadinma"
date: "14/09/2019"
output: html_document
---

```{r setup, include=FALSE}
library(pracma)
knitr::opts_chunk$set(echo = TRUE)
```

### (1) What is the rank of the matrix A?


```{r}

(A <- matrix(c(1, 2, 3, 4,-1, 0, 1, 3, 0, 1, -2, 1, 5, 4, -2, -3), nrow=4, byrow = TRUE))

```


```{r}

print(det(A))

```

#### The determinant of A is non-zero therefore, the rank of matrix A is 4 ()

```{r}

print(Rank(A))

```

### (2) Given an mxn matrix where m > n, what can be the maximum rank? The mini-
mum rank, assuming that the matrix is non-zero?

#### Answer:
#### Since the rank is the number of all non-zero row, the rank has to be no greater than the minimum value of either m or n.

### (3) What is the rank of matrix B?

```{r}

(B <- matrix(c(1, 2, 1, 3, 6, 3, 2, 4, 2), nrow=3, byrow = TRUE))

```
#### From the above matrix, Row 2 and Row 3 are multiples of Row 1 in varying degrees. Therefore,

```{r}
R1 <- 
R2 <- 
R3 <- 

a <- B[1, ]-(1/3)%*%B[2, ]
b <- B[3, ]-(2/3)%*%B[2, ]

matrix(c(a,b,B[2, ]), nrow = 3, byrow = T)

```

#### This implies that the rank of matirx B is 1

### 2. Problem set 2
#### Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

```{r}

(A <- matrix(c(1, 2, 3, 0, 4, 5, 0, 0, 6), nrow=3, byrow = T))

```

$$ Ax = \lambda x $$
$$ Ax - \lambda x = 0 $$

$$ det \left( A - \lambda I \right)x = 0$$

$$\begin{equation*}
\mathbf{} A-\lambda I = \left[\begin{matrix}
1 - \lambda & 2 & 4\\ 0 & 4-\lambda & 5\\ 0 & 0 & 6-\lambda
\end{matrix}\right] 
\end{equation*}
$$

The determinant of the matrix will be the product of the diagonal elements. Therefore,

$$det\left( A-\lambda I\right) = \left(1-\lambda\right)\left(4-\lambda \right)\left(6-\lambda \right) = 0 $$

$$ \left(4-\lambda - 4\lambda + \lambda^{2}\right)\left(6-\lambda \right) = 0$$


$$ \left(24-6\lambda - 24\lambda + 6\lambda^{2}-4\lambda + \lambda^{2} + 4\lambda^{2} - \lambda^3\right) = 0$$

#### The characteristic polynomial is:
$$-\lambda^3 + 11\lambda^2 - 34\lambda + 24 = 0$$

#### The eigen values are: 1, 4, and 6

At $\lambda_1 = 1$, the corresponding eigen vectors are:

$$\begin{equation*}
\mathbf{}\left[\begin{matrix}
1.0000\\ 0.0000\\ 0.0000
\end{matrix}\right] 
\end{equation*}
$$
At $\lambda_2 = 4$, the corresponding eigen vectors are:

$$\begin{equation*}
\mathbf{}\left[\begin{matrix}
1.6000\\ 2.5000\\ 1.0000
\end{matrix}\right] 
\end{equation*}
$$


At $\lambda_3 = 6$, the corresponding eigen vectors are:

$$\begin{equation*}
\mathbf{}\left[\begin{matrix}
1.6667\\ 1.0000\\ 0.0000
\end{matrix}\right] 
\end{equation*}
$$

























