---
title: "DATA605 Homework 13"
author: "Henry Otuadinma"
date: "24/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(stringr)
library(DT)

```

#### 1. Use integration by substitution to solve the integral below.
$\int{4e^{-7x}dx}$

solution
let  U = −7x

$dU = −7dx$

$dx = \frac{dU}{−7}$

$4\int{e^U\frac{dU}{−7}}$

$-\frac{4}{7}∫e^UdU$

$\frac{-4}{7}e^U+C$

$-\frac{4}{7}e^{−7x}+C$


#### 2.Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of 
$\frac{dN}{dt}=-\frac{3150}{t^4} - 220$ bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N(t) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter

solution


##### If we set N(1) = 6530, We can find The constant of integration, N0.

$\frac{dN}{dt}=-\frac{3150}{t^4} - 220$

$dN =(\frac{3150}{t^4}−220)dt$

$N=\int{\frac{3150}{t^4}dt} −\int{220dt}$

$N = N_0 − \frac{3150}{3t^3}− 220t$

$N(_1) = N_0 −\frac{1050}{1^3}−220(1)$

$N_0 = 6530 + 1050 + 220$

$N_0 = 7800$
 
$N = 7800−\frac{1050}{t^3}−220t$

#### 3. Find the total area of the red rectangles in the figure below, where the 3. equation of the line is f ( x ) = 2x - 9.

solution

##### Using a R function to find the difference between the areas under the curve

```{r}

fn = function(x) {2*x -9}
a <- integrate(fn, 4.5, 8.5)$value
(a <- round(as.numeric(a)))

```

#### Find the area of the region bounded by the graphs of the given equations.

$y = x2 - 2x - 2, y = x + 2$

solution

##### using R functions to find the difference between areas under the curve

```{r}

X = function(x) {x + 2}
Y = function(x) {x^2 -2*x -2}

```


```{r}
ax <- integrate(X, -1, 4)
yx <- integrate(Y, -1, 4)
(Area <- round((ax$value - yx$value),4))

```

#### A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

solution 

$f′(x) = 1.875−\frac{907.5}{x^2} = 0$

$1.875 − \frac{907.5}{x^2} = 0$

$1.875 = \frac{907.5}{x^2}$

$1.875* x^2 = 907.5$

$x^2 = \frac{907.5}{1.875}$

$x = \sqrt{\frac{907.5}{1.875}}$

$x = 22$


#### Use integration by parts to solve the integral below

$\int{ln(9x).x^6 dx}$

solution

$=\frac{1}{7}x^7 * ln(9x) − \int{\frac{1}{7}x^7 * \frac{1}{x}dx}$

$=\frac{1}{7}x^7 * ln(9x) - \int{\frac{1}{7}x^6dx}$

$= \frac{7}{49}x^7 * in(9x) - \frac{1}{49} x^7dx + C$

$= \frac{1}{49}x^7*(7ln(9x) - 1) + C$


##### 7. Determine whether f ( x ) is a probability density function on the interval 1, e6 . If not, determine the value of the definite integral.

$f(x) = \frac{1}{6x}$

solution

$F(x)=\int_1^{e^6}{f(x)dx}=1$ 

where $f(x) = \frac{1}{6x}$

=> $F(x) = \int{_1^{e^6}\frac{1}{6x}dx}$

$= \frac{1}{6}\int_1^{e^6}{\frac{1}{x}dx}$

$= \frac{1}{6}ln(x)|_1^{e^6}$

$= \frac{1}{6}[ln(e^6) - ln(1)]$

$= \frac{1}{6}[6-0]$

=> $F(x) = 1$

##### At interval $[1, e^6]$, the integral function F(x) resolves to 1. Therefore, F(x) is a probability density function at this interval.

































