---
title: "DATA 607 Week 1 Assignment"
author: "Henry Otuadinma"
date: "1 February 2019"
rpubs: "http://rpubs.com/jackv13/463542"
github: "https://github.com/henryvalentine/MSDS2019/raw/master/Classes/DATA%20607/Home%20works/DATA_607_Week_1_Assignment.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The mushroom dataset detalis

1. Title: Mushroom Database

2. Sources: 
    (a) Mushroom records drawn from The Audubon Society Field Guide to North
        American Mushrooms (1981). G. H. Lincoff (Pres.), New York: Alfred
        A. Knopf
    (b) Donor: Jeff Schlimmer (Jeffrey.Schlimmer@a.gp.cs.cmu.edu)
    (c) Date: 27 April 1987
    
This data set includes descriptions of hypothetical samples
    corresponding to 23 species of gilled mushrooms in the Agaricus and
    Lepiota Family (pp. 500-525).  Each species is identified as
    definitely edible, definitely poisonous, or of unknown edibility and
    not recommended.  This latter class was combined with the poisonous
    one.  The Guide clearly states that there is no simple rule for
    determining the edibility of a mushroom; no rule like ``leaflets
    three, let it be'' for Poisonous Oak and Ivy.

5. Number of Instances: 8124

6. Number of Attributes: 22 (all nominally valued)

7. Attribute Information: (classes: edible=e, poisonous=p)
     1. cap-shape:                bell=b,conical=c,convex=x,flat=f,
                                  knobbed=k,sunken=s
     2. cap-surface:              fibrous=f,grooves=g,scaly=y,smooth=s
     3. cap-color:                brown=n,buff=b,cinnamon=c,gray=g,green=r,
                                  pink=p,purple=u,red=e,white=w,yellow=y
     4. bruises?:                 bruises=t,no=f
     5. odor:                     almond=a,anise=l,creosote=c,fishy=y,foul=f,
                                  musty=m,none=n,pungent=p,spicy=s
     6. gill-attachment:          attached=a,descending=d,free=f,notched=n
     7. gill-spacing:             close=c,crowded=w,distant=d
     8. gill-size:                broad=b,narrow=n
     9. gill-color:               black=k,brown=n,buff=b,chocolate=h,gray=g,
                                  green=r,orange=o,pink=p,purple=u,red=e,
                                  white=w,yellow=y
    10. stalk-shape:              enlarging=e,tapering=t
    11. stalk-root:               bulbous=b,club=c,cup=u,equal=e,
                                  rhizomorphs=z,rooted=r,missing=?
    12. stalk-surface-above-ring: fibrous=f,scaly=y,silky=k,smooth=s
    13. stalk-surface-below-ring: fibrous=f,scaly=y,silky=k,smooth=s
    14. stalk-color-above-ring:   brown=n,buff=b,cinnamon=c,gray=g,orange=o,
                                  pink=p,red=e,white=w,yellow=y
    15. stalk-color-below-ring:   brown=n,buff=b,cinnamon=c,gray=g,orange=o,
                                  pink=p,red=e,white=w,yellow=y
    16. veil-type:                partial=p,universal=u
    17. veil-color:               brown=n,orange=o,white=w,yellow=y
    18. ring-number:              none=n,one=o,two=t
    19. ring-type:                cobwebby=c,evanescent=e,flaring=f,large=l,
                                  none=n,pendant=p,sheathing=s,zone=z
    20. spore-print-color:        black=k,brown=n,buff=b,chocolate=h,green=r,
                                  orange=o,purple=u,white=w,yellow=y
    21. population:               abundant=a,clustered=c,numerous=n,
                                  scattered=s,several=v,solitary=y
    22. habitat:                  grasses=g,leaves=l,meadows=m,paths=p,
                                  urban=u,waste=w,woods=d

8. Missing Attribute Values: 2480 of them (denoted by "?"), all for
   attribute #11.

9. Class Distribution: 
    --    edible: 4208 (51.8%)
    -- poisonous: 3916 (48.2%)
    --     total: 8124 instances

```{r}
download.file('https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data', 'agaricus-lepiota.data')
mushrooms <- read.table('agaricus-lepiota.data', sep=',', stringsAsFactors=FALSE)
```


```{r}
summary(mushrooms)

```

```{r}
head(mushrooms)

```

```{r}
dim(mushrooms)

```

```{r}

df <- subset(mushrooms, select = c(V1, V3, V17, V19))

names(df) <- c('class', 'cap-surface', 'veil-type', 'ring-number')

```

```{r}
head(df)
```

```{r}

## Renaming values in the columns

df$class[df$class=="p"]<- "poisonous"
df$class[df$class=="e"]<- "edible"

df['cap-surface'][df['cap-surface']=="f"]<- "fibrous"
df['cap-surface'][df['cap-surface']=="g"]<- "grooves"
df['cap-surface'][df['cap-surface']=="y"]<- "scaly"
df['cap-surface'][df['cap-surface']=="s"]<- "smooth"

df['veil-type'][df['veil-type']=="u"]<- "universal"
df['veil-type'][df['veil-type']=="p"]<- "partial"

df['ring-number'][df['ring-number']=="n"]<- "none"
df['ring-number'][df['ring-number']=="o"]<- "one"
df['ring-number'][df['ring-number']=="t"]<- "two"


```

```{r}

summary(df, 30)

```

```{r}

head(df, 30)

```

















