---
title: "DATA 607 Week 3 Homework"
author: "Henry Otuadinma"
date: "14 February 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```

#### 3. Copy the introductory example. The vector name stores the extracted names.

R> name
[1] "Moe Szyslak" "Burns, C. Montgomery" "Rev. Timothy Lovejoy"
[4] "Ned Flanders" "Simpson, Homer" "Dr. Julius Hibbert"

##### (a) Use the tools of this chapter to rearrange the vector so that all elements conform to 
##### the standard first_name last_name.

```{r}

rawData <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

```


```{r}

names <- unlist(str_extract_all(rawData, "[[:alpha:]., ]{2,}"))
names
```


```{r}
# remove titles from the names
firstLastName <- str_replace_all(names, "\\b\\w{2,}?\\.", '') # the '\\b' is not necessary but can be useful 
# when the title is preceeded by another word in a sentence
firstLastName
```

##### (b) Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).

```{r}

data.frame(names = names, hasTitle = str_detect(names, "\\b\\w{2,}?\\."))

```

##### (c) Construct a logical vector indicating whether a character has a second name.

```{r}

data.frame(names = firstLastName, hasSecondName = str_detect(names, "\\b\\w{1,}?\\."))

```

#### 4. Describe the types of strings that conform to the following regular expressions and
#### construct an example that is matched by the regular expression.

(a) [0-9]+\\$

```{r}

values <- c('The shoe worths $100', 'ab988uy$', '1234', '5000$')
matcher <- '[0-9]+\\$'

data.frame(value = values, matched = str_detect(values, matcher))

```

(b) \\b[a-z]{1,4}\\b

```{r}

values <- c('The shoe worths $100', 'My Name Is Subzero', 'My name Is subzero', '5000$')
matcher <- '\\b[a-z]{1,4}\\b'

data.frame(value = values, matched = str_detect(values, matcher))

```

(c) .*?\\.txt$

```{r}

values <- c('chess_players_ratings.txt', 'chess_players_ratings.pdf', 'Chess_PLAYERS_raTings.txt', 'Chess_PlAYeRS.tx')
matcher <- '.*?\\.txt$'

data.frame(value = values, matched = str_detect(values, matcher))

```

(d) \\d{2}/\\d{2}/\\d{4}

```{r}

values <- c('3/50', '20/50/5504', '99/100-33/50', '11/22/4840')
matcher <- '\\d{2}/\\d{2}/\\d{4}'

data.frame(value = values, matched = str_detect(values, matcher))
```


(e) <(.+?)>.+?</\\1>

```{r}

values <- c('<a>click here to sign up</a>', 'sps.cuny.edu', '<button> a classic button</button>', 'https://www.amazon.com')
matcher <- '<(.+?)>.+?</\\1>'

data.frame(value = values, matched = str_detect(values, matcher))
```


#### 9. The following code hides a secret message. 
Crack it with R and regular expressions.
##### Hint: Some of the characters are more revealing than others! 
The code snippet is also available in the materials at www.r-datacollection.com.

```{r}

#With close observations, the hidden message can be seen to be a meaningful word in upper cases mixed among the lot

nerdMs <- 'clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr'

superNerd <- str_replace_all(str_replace_all(nerdMs,"([a-z])|([0-9])|\n",""), "\\.", " ")

superNerd
```
















