---
title: "DATA605 Discussion 13"
author: "Henry Otuadinma"
date: "21/11/2019"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Find the total area enclosed by the functions f and g:

$f(x) = 2x^2 + 5x - 31$, $ g(x) = x^2 + 4x - 1$

Solution

##### using R to put up functions f and g for the computations:

```{r}

f <- function(x) {2*x^2+5*x-3}
g <- function(x) {x^2+4*x-1}

```


```{r}
curve(f, -3, 3)
curve(g, -3, 3, add=TRUE)

```

##### Looking for the value where the point of intersection falls below 0:

```{r}

(rootF <- uniroot(function(x)  f(x) - g(x)  , c(-5000,-0.01), tol=1e-8)) 

```

##### What is the root of the function:

```{r}

rootF$root

```

##### Looking for the value where the point of intersection is above 0:

```{r}

(root0 <- uniroot(function(x)  f(x) - g(x)  , c(0.0001, 5000), tol=1e-8))
 
```

##### What is the root of the function:

```{r}

root0$root

```

#### Therefore, the area between the curves will be g minus f at x(-2, 1)



##### integrating for g
```{r}

(areaG <- integrate(g, lower = -2, upper = 1))

```


##### integrating for f

```{r}

(areaF <- integrate(f, lower = -2, upper = 1))

```


##### compute difference between g_area and f_area

```{r}

areaG$value - areaF$value

```








