---
title: "DATA 607 Week 1 Lab"
author: "Henry Otuadinma"
date: "31 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rvest)
```

## You should first study the data and its associated description (i.e. "data dictionary"). 
## You should take the data, and create an R dataframe with a subset of the columns in the dataset. 
## You should provide relevant column names. Your deliverable is the R Markdown code to perform 
## these transformation tasks.

### Solutions:

The given link in the Hands on Lab pdf file could not work so I was able to google around and found the dataset on the link below:
https://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version1

```{r}
pittsburgBridges <- read.csv('https://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version1', header=TRUE,na.strings = "?")
```

```{r}
summary(pittsburgBridges)

```




```{r}
head(pittsburgBridges)

```

```{r}
dim(pittsburgBridges)

```

### The complete details about this dataset including the column names and contributors can be gotten from the url below:
ftp://ftp.ics.uci.edu/pub/machine-learning-databases/bridges/bridges.names

```{r}
textStream <- readLines('ftp://ftp.ics.uci.edu/pub/machine-learning-databases/bridges/bridges.names')

```

```{r}
#replace tabs with spaces, trim whit spaces at beginning and end, and remove quotations
print(gsub("\t"," ",trimws(textStream)), quote=FALSE)
```

### From the above information, I can conveniently rename the columns as follows

```{r}
names(pittsburgBridges) <- c('IDENTIF','RIVER', 'LOCATION','ERECTED','PURPOSE','LENGTH','LANES','CLEAR-G','T-OR-D','MATERIAL','SPAN','REL-L','TYPE')

```

```{r}
head(pittsburgBridges)
```

### A subset of the coumns can be used to create a new dataframe as follows:

```{r}
newBridges <- pittsburgBridges[, c(1,2,3,4,7,9)]
```

```{r}
head(newBridges)
```